Making Deep Q-learning Methods Robust to Time Discretization

### 2019-01-19 paper/icml_drau/advup.tex

As seen above, there is no continuous time limit to Q-learning. In
continuous time, $Q^\pi$ is independent of actions and can thus not be
used to select promising actions.  In near-continuous time,
$Q^\pi_\deltat$ still depends on actions, and could still be used to
choose promising actions under the current policy. However, when
approximating $Q^\pi_\deltat$, if the approximation error is much bigger
than $\bigO(\deltat)$, approximation errors dominates, and the ranking of
actions given by the approximation is likely to be erroneous.

% When the approximate Q-function is initialized, if the effect of
% actions on the Q-function is order of magnitudes higher than what it should be,
% approximating $Q^\pi_\deltat$ is likely to be difficult.
%Besides, the error is likely
%to be further propagated when the equation used to update our approximate $Q$
%function relies on bootstraping, as is the case for usual temporal difference
%derived methods.

\subsection{Advantage Updating}
\label{subsec:reparam}
To define an object which contains the same information on actions as
$Q^\pi_\deltat$, but admits an action dependant continuous time limit, it is
natural to define \cite{adv_upd}
\begin{align}
	A^\pi_\deltat(s, a) &\mathdef \frac{Q^\pi_\deltat(s,a) - V^\pi_\deltat(s)}{\deltat},
    \label{eq:adv}
\end{align}
a rescaled version of the advantage function, as the difference between between
$Q^\pi_\deltat(s, a)$ and $V^\pi_\deltat(s)$ is of order $\bigO(\deltat)$.
Contrary to the action state value function, the rescaled advantage function converges when $\deltat$ goes to $0$
to an action dependent quantity which contains the necessary information for policy improvement:

@@

The discretized Q-function can be rewritten as
\begin{equation}
	Q^\pi_\deltat(s, a) = V^\pi_\deltat(s) + \deltat A^\pi_\deltat(s, a).
	\label{eq:reparam_q_pi}
\end{equation}
A natural way of approximating $V^\pi_\deltat$ and $A^\pi_\deltat$ is to apply
\emph{Sarsa} or Q-learning to a reparameterized Q-function approximator
\begin{equation}
	Q_\Theta(s, a) = V_{\theta}(s) + \deltat A_{\psi}(s, a).
\end{equation}
with $\Theta = (\theta, \pi)$. To avoid cumbersome notations, in what follows,
parameter indices are dropped.  At initialization, if both $V_{\theta}$ and
$A_{\psi}$ are initialized independantly of $\deltat$, this parameterization
reasonably scales the contribution of actions relative to states in the value
of $Q$.  Our goal is for $V_\theta$ to approximate $V^\pi_\deltat$ and for
$A_{\psi}$ to approximate $A^\pi_\deltat$.

@@

To learn $V_{\theta}$ and $A_{\psi}$, variants of Q-learning for continuous and
discrete action spaces are used. In details, $V_{\theta}$ and $A_{\psi}$ are learnt
using approximate version of the following bellman equation, which holds for the true $A^\pi_\deltat$
and $V^\pi_\deltat$, with a greedy exploitation policy $\pi(s) = \text{argmax}_{a'}A^\pi_\deltat(s, a')$
\begin{align}
	V^{\pi}_\deltat(s) + \deltat A^{\pi}_\deltat(s, a) &= r\deltat + \gamma^{\deltat}  \E_{s'} V^{\pi}_\deltat(s')\label{eq:bellman_A}\\
	A^{\pi}_\deltat(s, \pi(s)) &= 0\label{eq:max_A}.
\end{align}
Maximization over actions is implemented exactly in the discrete actions case,
and approximated by a policy neural network $\pi_\phi(s)$, trained to maximize $A(s,
\pi_\phi(s))$ in the continuous case.  This is similar to what is done in
\cite{ddpg}.

>>>>

% When the approximate $Q$-function is initialized, if the effect of
% actions on the $Q$-function is order of magnitudes higher than what it should be,
% approximating $Q^\pi_\deltat$ is likely to be difficult.
%Besides, the error is likely
%to be further propagated when the equation used to update our approximate $Q$
%function relies on bootstraping, as is the case for usual temporal difference
%derived methods.

\subsection{Advantage Updating}
\label{subsec:reparam}

As seen above, there is no continuous time limit to $Q$-learning, because
$Q^\pi$ becomes independent of actions and thus cannot be
used to select actions.  In near-continuous time,
$Q^\pi_\deltat$ still depends on actions, and could still be used to
choose actions. However, when
approximating $Q^\pi_\deltat$, if the approximation error is much larger
than $\bigO(\deltat)$, this error dominates, the ranking of
actions given by the approximated $Q^\pi_\deltat$ is likely to be erroneous.

To define an object which contains the same information on actions as
$Q^\pi_\deltat$, but admits a learnable action-dependent limit, it is
natural to define \cite{adv_upd}
\begin{align}
	A^\pi_\deltat(s, a) &\mathdef \frac{Q^\pi_\deltat(s,a) - V^\pi_\deltat(s)}{\deltat},
    \label{eq:adv}
\end{align}
a rescaled version of the advantage function, as the difference between between
$Q^\pi_\deltat(s, a)$ and $V^\pi_\deltat(s)$ is of order
$\bigO(\deltat)$. This amounts to splitting $Q$ into value and advantage,
and observing that these scale very differently when $\deltat\to 0$.
%Contrary to the action state value function,
This
rescaled advantage function converges when $\deltat\to 0$
to an action-dependent quantity which, if it can be learned, contains the necessary information for policy improvement:

@@

The discretized $Q$-function rewrites as
\begin{equation}
	Q^\pi_\deltat(s, a) = V^\pi_\deltat(s) + \deltat A^\pi_\deltat(s, a).
	\label{eq:reparam_q_pi}
\end{equation}
A natural way of approximating $V^\pi_\deltat$ and $A^\pi_\deltat$ is to apply
\emph{Sarsa} or $Q$-learning to a reparameterized $Q$-function approximator
\begin{equation}
	Q_\Theta(s, a) = V_{\theta}(s) + \deltat A_{\psi}(s, a).
\end{equation}
with $\Theta = (\theta, \pi)$. To avoid cumbersome notations, in what follows,
parameter indices are dropped.  At initialization, if both $V_{\theta}$ and
$A_{\psi}$ are initialized independantly of $\deltat$, this parameterization
reasonably scales the contribution of actions relative to states in the value
of $Q$.  Our goal is for $V_\theta$ to approximate $V^\pi_\deltat$ and for
$A_{\psi}$ to approximate $A^\pi_\deltat$.

@@

To learn $V_{\theta}$ and $A_{\psi}$, variants of $Q$-learning for continuous and
discrete action spaces are used. In details, $V_{\theta}$ and $A_{\psi}$ are learnt
using approximate version of the following bellman equation, which holds for the true $A^\pi_\deltat$
and $V^\pi_\deltat$, with a greedy exploitation policy $\pi(s) = \text{argmax}_{a'}A^\pi_\deltat(s, a')$
\begin{align}
	V^{\pi}_\deltat(s) + \deltat A^{\pi}_\deltat(s, a) &= r\deltat + \gamma^{\deltat}  \E_{s'} V^{\pi}_\deltat(s')\label{eq:bellman_A}\\
	A^{\pi}_\deltat(s, \pi(s)) &= 0\label{eq:max_A}.
\end{align}
Maximization over actions is implemented exactly in the discrete actions case,
and approximated by a policy neural network $\pi_\phi(s)$, trained to maximize $A(s,
\pi_\phi(s))$ in the continuous case.  This is similar to what is done in
\cite{ddpg}.

### 2019-01-19 paper/icml_drau/intro.tex

Standard approaches based on estimation of state-action
value functions, such as
\emph{Deep $Q$-learning} (DQN,~\citealt{dqn}) and \emph{Deep deterministic policy
gradient} (DDPG,~\citealt{ddpg}) are not at all \TODO{Remove familiar?}robust to changes in time discretization. This is shown experimentally in Sec.~\ref{sec:exp}. 
Intuitively, as the discretization timestep
decreases, the effect of individual actions on the total return decreases
too: $Q^*(s,a)$ is the value of playing action $a$ then playing optimally,
and if $a$ is only maintained for a very short time its advantage over
other actions will be accordingly small. (This continuity effect occurs even with a
well-adjusted decay factor $\gamma$.) \TODO{Really relevant in the introduction?}

>>>>

Standard approaches based on estimation of state-action
value functions, such as
\emph{Deep $Q$-learning} (DQN,~\citealt{dqn}) and \emph{Deep deterministic policy
gradient} (DDPG,~\citealt{ddpg}) are not at all \TODO{Remove familiar?
NDY it's not particularly familiar}robust to changes in time discretization. This is shown experimentally in Sec.~\ref{sec:exp}. 
Intuitively, as the discretization timestep
decreases, the effect of individual actions on the total return decreases
too: $Q^*(s,a)$ is the value of playing action $a$ then playing optimally,
and if $a$ is only maintained for a very short time its advantage over
other actions will be accordingly small. (This continuity effect occurs even with a
well-adjusted decay factor $\gamma$.) \TODO{Really relevant in the introduction?}
